Type Error Telemetry at Scale: Pseudonymized and Stochiastic

### 2023-09-20 programming23/paper.tex
outline rest of Data section

\begin{table}
  \caption{Selected Error Labels}
  %% from type analysis, but they're not all really type errors
  %% TODO review selection after analyzing the full data
  \label{t:type-error-labels}

@@

@@

\FILL{} compare \mnonstrict{} and \mstrict{}.
Can't compare to \mnocheck{} with this table, though, so maybe it make sense to
put this question later.

@@

\begin{itemize}
  \item
    As we noted above, the most popular errors are syntax errors
    (\pct{FILL} \mnonstrict{}, \pct{FILL} \mstrict{})..
    This is undoubtedly due to the stochiastic nature
    of our telemetry, which has no way to avoid
    reporting edits that happen in the middle of things.
    For example, a creator who writes a method call
    \code{bucket.countFish()} letter-by-letter will
    get type analysis to find errors like \code{UnknownSymbol(buck)}
    and \code{UnknownProperty(bucket.c)}.
    These are transient issues, but telemetry may catch them.

@@

  \item
    \FILL{} which errors that NEVER showed up?
    Nothing for \code{CodeTooComplex} or \code{UnificationTooComplex} is great news.
    But what about the others? FILL

@@

\subsection{NEXT}

Second,
the stronger analysis modes are roughly 10x less popular than weaker ones
(across the \pct{99} of sessions that stick with one mode):

\pct{90} of sessions use nocheck mode
\pct{9.8} use nonstrict mode
\pct{0.2} use strict mode

Third,
In the edit range, we know exactly which errors showed up. This leads to a few
takeaways, such as:

Only 3 sessions (out of +300K) hit the typechecker's internal limits on problem
size --- CodeTooComplex errors. Good!

About \pct{90} of type errors are probably due to typos (SyntaxError, UnknownSymbol,
UnknownProperty). These are not propoer type errors.

10 errors never showed up. These include DynamicPropertyLookupOnClassesUnsafe and
SwappedGenericTypeParameter

Fourth,
Codebase size and session length tend to follow a power law. 
One aspect of size for nonstrict sessions: the number of
lines in the edit range. Over 80 sessions have 20-line ranges, about 40 have a
30-line range, etc.

@@

\begin{figure}[t]\centering
% TODO diff vs 1st, not previous

  TODO bars instead, percent only, don't care about counts

  \begin{tabular}{lr@{}rr@{}rr@{}r}
    & \zerowidth{Add} & & \zerowidth{Keep} & & \zerowidth{Drop} \\\midrule
    \mnocheck{} & 48378 & [\pct{46.83}] & 9440 & [\pct{9.14}] & 45479 & [\pct{44.03}] \\
    \mnonstrict{} & 19491 & [\pct{39.63}] & 9567 & [\pct{19.45}] & 20121 & [\pct{40.91}] \\
    \mstrict{} & 733 & [\pct{39.18}] & 368 & [\pct{19.67}] & 770 & [\pct{41.15}] \\
  \end{tabular}
  \caption{FILL error density}
  \label{f:error-delta}
\end{figure}

@@

@@

\section{Interpretation}

\subsection{Internal Error}
%% out/te-editrange-*

@@

\subsection{TBD}

Error 1000 = type mismatch (subtyping etc);
1001 = unknown symbol;
\ldots

% https://github.com/Roblox/luau/blob/master/Analysis/include/Luau/Error.h

@@

\subsection{Module Switches}

FILL why study, what implications?

FILL need \%s out of all module switches.

How many module switches exit a module with errors?
\mnocheck{} 24, \mnonstrict{} 9, \mstrict{} 3.
Rare?
More common for forcedstrict:
 \mnocheck{} 190, \mnonstrict{} 16, \mstrict{} 2.

How many module switches go to a module with errors?
\mnocheck{} 70, nonstrict 9, strict 4
For FS: \mnocheck{} 400, \mnonstrict{} 39, \mstrict{} 6.

\subsection{Types and Quality of Life}

FILL error density: number vs range size

TODO nocheck accumulated errors?

\subsection{Example Sessions}

FILL example sessions.

\subsection{Asset Creation vs Scripting}

Data model. Point to discussion above.

Big fluctuations in errors.
FS disagrees with TE.
Probably data model!

\section{Threats to Validity}
\label{s:threats}

@@

  \begin{tabular}{l@{~}cccccc}
    &             & PII       & Session ID & Deterministic & Private Data \\\midrule
    & Transparent & \chkNo    & \chkNo     & \chkYes       & \chkNo      \\
  * & Us          & \chkNo    & \chkYes    & \chkNo        & \chkYes     \\
    & \code{.NET} & \chkMaybe & \chkYes    & \chkYes       & \chkYes     \\
    & VS Code     & \chkYes   & \chkYes    & \chkYes       & \chkYes     \\
  \end{tabular}
\end{table}

@@

\section{Discussion}
\label{s:conclusion}
\label{s:discussion}

Findings:

@@

FILL regrets, lessons learned, advice for next time

@@

Thanks to Benjamin Chung for feedback on plots and data analysis advice.
Greenman was supported by
grant \href{https://nsf.gov/awardsearch/showAward?AWD_ID=2030859&HistoricalAwards=false}{2030859}
to the CRA for the \href{https://cifellows2020.org}{CIFellows} project.

>>>>

\begin{table}
  \caption{Selected Error Labels}
  % https://github.com/Roblox/luau/blob/master/Analysis/include/Luau/Error.h
  %% from type analysis, but they're not all really type errors
  %% TODO review selection after analyzing the full data
  \label{t:type-error-labels}

@@

%% \paragraph{Module Switches vs Errors}
%% FILL need \%s out of all module switches.
%% How many module switches exit a module with errors?
%% > OLD DATA
%% > \mnocheck{} 24, \mnonstrict{} 9, \mstrict{} 3.
%% > Rare?
%% > More common for forcedstrict:
%% >  \mnocheck{} 190, \mnonstrict{} 16, \mstrict{} 2.
%% >
%% > How many module switches go to a module with errors?
%% > \mnocheck{} 70, nonstrict 9, strict 4
%% > For FS: \mnocheck{} 400, \mnonstrict{} 39, \mstrict{} 6.

@@

@@

\begin{itemize}
  \item
    As we noted above, the most popular errors are syntax errors
    (\FILL{} \pct{0} \mnonstrict{}, \pct{0} \mstrict{})..
    This is undoubtedly due to the stochiastic nature
    of our telemetry, which has no way to avoid
    reporting edits that happen in the middle of things.
    For example, a creator who writes a method call
    \code{bucket.countFish()} letter-by-letter will
    get type analysis to find errors like \code{UnknownSymbol(buck)}
    and \code{UnknownProperty(bucket.c)}.
    These are transient issues, but telemetry may catch them.

@@

  \item
    \FILL{} which errors that NEVER showed up?
    Nothing for \code{CodeTooComplex} or \code{UnificationTooComplex} is great news.
    But what about the others?
    \code{DynamicPropertyLookupOnClassesUnsafe}?
    \code{SwappedGenericTypeParameter}?

@@

\subsection{Forced Strict Errors}

\FILL{} I (ben) think we need to talk more about forced strict errors here, or soon.
Otherwise people will forget they exist in the data.
(But if they're not useful to us, remove them!)

\subsection{Type Error Density}

\begin{figure}[t]\centering
  % TODO try diff vs first record instead of previous?

  % TODO use bars instead, percent only, don't care about counts
  \begin{tabular}{lr@{}rr@{}rr@{}r}
    & \zerowidth{Add} & & \zerowidth{Keep} & & \zerowidth{Drop} \\\midrule
    \mnocheck{} & 48378 & [\pct{46.83}] & 9440 & [\pct{9.14}] & 45479 & [\pct{44.03}] \\
    \mnonstrict{} & 19491 & [\pct{39.63}] & 9567 & [\pct{19.45}] & 20121 & [\pct{40.91}] \\
    \mstrict{} & 733 & [\pct{39.18}] & 368 & [\pct{19.67}] & 770 & [\pct{41.15}] \\
  \end{tabular}
  \caption{FILL error density}
  \label{f:error-delta}
\end{figure}

@@

\subsection{Data Model}

% -\subsection{Asset Creation vs Scripting}
% -
% -Data model. Point to discussion above.
% -
% -Big fluctuations in errors.
% -FS disagrees with TE.
% -Probably data model!

@@

\subsection{Timeline}

@@

\section{Discussion}
\label{s:discussion}

\FILL{} good things first,
then regrets about the noisy stochiastic telemetry,
then lessons?
then threats.

\paragraph{Good: No Internal Error}

@@

\FILL{} compare \mnonstrict{} and \mstrict{}.
Use edit range~(\cref{t:type-error-survival} for NS vs S)
and error density~(\cref{f:error-delta})

TODO nocheck accumulated fs errors?

\paragraph{Bad: Noisy Data}

@@

\subsection{Threats to Validity}
\label{s:threats}

@@

  \begin{tabular}{l@{~}cccccc}
    &             & PII       & Session ID & Deterministic & Private Data (???) \\\midrule
    & Transparent & \chkNo    & \chkNo     & \chkYes       & \chkNo      \\
  * & Us          & \chkNo    & \chkYes    & \chkNo        & \chkYes     \\
    & Diff.Priv   & $\epsilon$ & \chkYes   & \chkYes       & \chkYes     \\
    & Liblit      & ?         & ?          & ?             & ?           \\
    & \code{.NET} & \chkMaybe & \chkYes    & \chkYes       & \chkYes     \\
    & VS Code     & \chkYes   & \chkYes    & \chkYes       & \chkYes     \\
  \end{tabular}
\end{table}

@@

\section{Conclusion}
\label{s:conclusion}

Quickly summarize findings:

@@

Quick regrets, lessons learned, advice for next time

@@

Thanks to Benjamin Chung for several helpful discussions about data analysis
and effective plotting.
Greenman was supported by
grant \href{https://nsf.gov/awardsearch/showAward?AWD_ID=2030859&HistoricalAwards=false}{2030859}
to the CRA for the \href{https://cifellows2020.org}{CIFellows} project.
